---
title: "ETC5523 Assignment 2"
author: Travis Rutledge 
date: 27/08/2024 
format: 
  html: default
  pdf: default
theme:
  light: flatly
  dark: darkly

---

```{r setup, include = TRUE, echo = FALSE, cache = FALSE}

# .qmd settings
knitr::opts_chunk$set(
  echo = TRUE,
  eval = TRUE,
  out.width = "70%",
  fig.width = 8, 
  fig.height = 6,
  fig.retina = 3)
```



```{r libraries, message = FALSE, echo=FALSE, warning=FALSE, eval=TRUE}
library(tidyverse)
library(readr)
library(hrbrthemes)
library(dplyr)
library(ggplot2)
library(here)
library(knitr) 
```

```{r load_dataset, message = FALSE, echo=FALSE, warning=FALSE, eval=TRUE}
medals <- read_csv(here::here("data", "medals.csv"))
```

# EXECUTIVE SUMMARY
Our team is proposing a new way to measure Olympic performance called Weighted Total Score, which gives a country 3 points for a winning a gold medal, 2 points for a silver, and 1 point for a bronze. The benefit of this measure is that it is inclusive of all medals won while scaling based type of medal. A limitation of this measure is the potential difficulty of the public adopting a new measure. To help with this, the new Weighted Total measure can be visualised alongside medal counts to provide clarity and familiarity. 


# INTRODUCTION TO THE PROBLEM 
The current approach to measuring a country's Olympic performance is counting gold medals. In other words, whichever country wins the most gold medals is determined to be the top performer for the games. While silver and bronze medals are recorded and visualised, they are not considered unless two countries are tied on medal count. And, although this approach is great in its' simplicity, it does have its problems. 

There are two problems with the IOC's current approach to counting gold medals and only looking at silver and bronze when there is a tie:

1. Athletes who won silver and bronze medals do not have their medals counted towards their country. They may feel personal success for reprsenting their country and winning a medal, but their achievement is not counted towards the official tally unless there is a tie for gold medals. 
2. Some argue that only tallying gold medals and not other medals is not a comprehensive look at the performance of a country. A country may win dozens of silver and bronze medals and no gold medals, and they would be ranked as one of the weakest country's in the games. Some countries like the United States and their media outlets report all medals earned, as they view it important to tally both gold medals own and total medals won[source](https://www.theguardian.com/sport/article/2024/aug/10/paris-olympics-2024-how-does-medal-tally-work-order). But, this is not how the IOC tallies results.

Others have proposed another way to ranking countries performance at the Olympics. Researchers Robert C Duncan and Andrew Parece published their own ranking method in the Journal of Sports Analytics that accounts for a country's probability of winning a medal if populations were all the same across countries. This approach was dubbed the "Duncan-Parece model". Although this calculation is useful in addressing population, the downside is that it may be difficult for the public to both understand and trust a calculation that is based on probability and weighting.

# KEY RESULTS 
## Introduction of a new measure - weighted medal score 
The goal is to create a new measure of tallying Olympic results per country that accomplishes three things:

1. It is inclusive of all medals won 
2. It prioritises gold medals won
3. It is simple to understand by the public

My proposed calculation is called the Weighted Medal Score (coloquially dubbed "Olympic Points"). The calculation gives 3 points to gold medals, 2 points to silver medals, and 1 point to bronze medals and then totals the scores for each country. The final score for each country is not the total count of all medals. Instead, it's a weighted score showing how productive a country is in winning medals.

The benefits of this new calculation is that it _______.

The biggest risk or limtiation of this measure is that it requires the public to shift their previous understanding of tallying countries' performance from medals to points. The public may be confused at first as to what a point or score means, and they might assume that the score is simply a medal count. To address this risk, the Weighted Medal Score will be visualised alongside medal counts. That way, the public will have a better opportunity to learn what the calculation is, how it is measured, and they can still refer to the previous way to tally results by looking at the medal counts. 



## Figure 

```{r bar_chart}
medals_with_score <- medals %>%
  mutate(Score = (G*3) + (S*2) + (B*1))
        
gold_bar_chart <- medals_with_score %>%
  filter(G >10) %>%
  mutate(Country = factor(Country, levels = Country[order(G, decreasing = FALSE)])) %>%
  ggplot(aes(x = G, y = Country)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Score),
            hjust = -.2,
            color = "gold",
            size = 3.5) +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.x = element_blank()) +
  labs(title = "China and the United States won the most gold medals at the 2024 Olympics")

bar_chart <- medals_with_score %>%
  filter(Score > 50) %>%
  mutate(Country = factor(Country, levels = Country[order(Score, decreasing = FALSE)])) %>%
  ggplot(aes(x = Score, y = Country)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Score),
            hjust = -.2,
            color = "grey",
            size = 3.5) +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.x = element_blank()) +
  labs(title = "The United States leads in the Medal Count Score",
       subtitle = "Gold medal = 3 points, silver medal = 2 points, bronze = 1 point")
```


```{r variable_names, messsage = FALSE, echo=FALSE, warning=FALSE, eval=FALSE}
#| label: tbl-1
#| tbl-cap: "Variable Names of the LGBTQ+ Dataset"
kable(data[25:30,])
```

# METHODS 
## Data Source 

## Data Dictionary 

# CONCLUSION 

